---
title: "MillerHW12"
author: "Larri Miller"
date: "11/23/2021"
output: html_document
---

I got tired of using .r files, so I'm using markdown instead to make this cleaner.

Pulling in replication data:
```{r}
q1 <- read.delim("LariReplicationResQuestOneProcVar1.txt", sep=",")
q2 <- read.delim("LariReplicationResQuestOneProcVar2.txt", sep=",")
q3 <- read.delim("LariReplicationResQuestOneProcVar3.txt", sep=",")
q4 <- read.delim("LariReplicationResQuestOneProcVar4.txt", sep=",")
q5 <- read.delim("LariReplicationResQuestTwoProcVar1.txt", sep=",")
q6 <- read.delim("LariReplicationResQuestTwoProcVar2.txt", sep=",")
q7 <- read.delim("LariReplicationResQuestTwoProcVar3.txt", sep=",")
q8 <- read.delim("LariReplicationResQuestTwoProcVar4.txt", sep=",")
```

t value code: 
```{r}
tval1=function(d1, mu=0.5){
  as.numeric(t.test(d1,mu=0.5)$stat)
}

tval2=function(d1, d2){
  as.numeric(t.test(d1, d2, paired=TRUE)$stat)
}
```

Bayes Factor Code
```{r}
prior=function(es,mu=0,sigma=1) dnorm(es,mu,sigma)

condPred=function(t,es,N){
  df=N-1
  ncp=es*sqrt(N)
  cpred=suppressWarnings(dt(t,df,ncp))
  return(cpred)
}

joint=function(t,es,N,mu=1,sigma=0) prior(es,mu,sigma)*condPred(t,es,N)
  
marPred=function(t,N,mu=0,sigma=1){
  mpred=rep(NA,length(t))
  for(i in 1:length(t)){
    mpred[i]=integrate(joint,lower=-Inf,upper=Inf,
                       t=t[i],N=N,mu=mu,sigma=sigma)$val
  }
  return(mpred)
}


nullPred=function(t,N) dt(t,N-1)
```

Function to calculate posterior probability of an effect
```{r}
postProb = function(bayesFactor){
  bayesFactor/(bayesFactor+1)
}
```

Approx post & see inference
``` {r}
approxPost=function(tObs,nObs,muPri=0,sigmaPri=1){
  #par(mfcol=c(1,1),mar=c(4.5,4.5,1,1))
  
  normCons=marPred(tObs,nObs,muPri,sigmaPri)
  
  posterior=function(es) joint(tObs,es,nObs,muPri,sigmaPri)/
    normCons
  
  esSeq=seq(-4,4,length.out=1000)
  post=posterior(esSeq)
  
  mode=esSeq[post==max(post)][1]
  
  rats=post/max(post)
  ratMiss=abs(rats-(dnorm(1)/dnorm(0)))
  oneSdAbove=esSeq[esSeq>mode & 
                     ratMiss==min(ratMiss[esSeq>mode])]
  oneSdBelow=esSeq[esSeq<mode & 
                     ratMiss==min(ratMiss[esSeq<mode])]
  
  devs=abs(c(oneSdAbove,oneSdBelow) - mode)
  
  muPost=mode
  sigmaPost=mean(devs)
  
  # plot(esSeq,post,type='l')
  # points(esSeq,dnorm(esSeq,muPost,sigmaPost),
  #        type="l",col="blue",lty=2)
  
  output=c(muPost,sigmaPost)
  return(output)
}

seeInference=function(tObs,nObs,muPri=0,sigmaPri=1){
  par(mfcol=c(3,1),mar=c(2.5,4.5,1,4.5))
  lineWid=1.5 #just sets line width for plots
  
  normCons=marPred(tObs,nObs,muPri,sigmaPri)
  
  posterior=function(es) joint(tObs,es,nObs,muPri,sigmaPri)/
    normCons
  
  esSeq=seq(-4,4,length.out=100)
  
  #plot prior
  dens=prior(esSeq,muPri,sigmaPri)
  yMax=max(dens)*1.2
  plot(esSeq,dens,type='l',
       bty='n',ylab="Prior",
       ylim=c(0,yMax),las=1,
       xlab="",lwd=lineWid)
  abline(h=0)
  
  #plot likelihood
  lik=condPred(tObs,esSeq,nObs)
  yMax=max(lik)*1.2
  plot(esSeq,lik,type='l',
       bty='n',ylab="Likelihood",
       ylim=c(0,yMax),las=1,
       xlab="",lwd=lineWid)
  abline(h=0)
  
  #get a Gaussian that approximates
  #  the posterior distribution
  distPost=approxPost(tObs,nObs,muPri,sigmaPri)
  muPost=distPost[1]
  sigmaPost=distPost[2]
  
  #plot posterior  
  post=posterior(esSeq)
  yMax=max(post)*1.2
  plot(esSeq,post,type='l',
       bty='n',ylab="Posterior",
       ylim=c(0,yMax),las=1,
       xlab="",lwd=lineWid)
  abline(h=0)
  
  points(esSeq,dnorm(esSeq,muPost,sigmaPost),
         type='l',col='orange',lty=3,lwd=lineWid)
  
  return(normCons)
}
```


Q1
Bayes Factor: 1.809163
Post Prob: 0.6440221
Mu: 0.7007007 Sigma: 0.1601602
```{r}
# Bayes Factor
tval1(q1$PropCor)
marPred(2.037173, 20, 0.925, 0.228)/nullPred(2.037173, 20)

# Post Prob
postProb(1.809163)

# distribution
seeInference(2.037173, 20, 0.925, 0.228)
approxPost(2.037173, 20, 0.925, 0.228)
```

Q2
Bayes Factor: 29271.95
Post Prob: 0.9999658
Mu: 1.2692693 Sigma: 0.1961962
```{r}
# Bayes Factor
tval1(q2$PropCor)
marPred(6.167072, 20, 1.197, 0.252)/nullPred(6.167072, 20)

# Post Prob
postProb(29271.95)

# distribution
seeInference(6.167072, 20, 1.197, 0.252)
approxPost(6.167072, 20, 1.197, 0.252)
```

Q3
Bayes Factor: 0.3493036
Post Prob: 0.2588769
Mu: 0.4684685 Sigma: 0.1121121
```{r}
# Bayes Factor
tval1(q3$PropCor)
marPred(0.734171, 20, 0.565, 0.128)/nullPred(0.734171, 20)

# Post Prob
postProb(0.3493036)

# distribution
seeInference(0.734171, 20, 0.565, 0.128)
approxPost(0.734171, 20, 0.565, 0.128)
```

Q4
Bayes Factor: 0.03165432
Post Prob: 0.03068307
Mu: 0.3323323 Sigma: 0.1081081
```{r}
# Bayes Factor
tval1(q4$PropCor)
marPred(-0.9266915, 20, 0.501, 0.128)/nullPred(-0.9266915, 20)

# Post prob
postProb(0.03165432)

# disitrbution
seeInference(-0.9266915, 20, 0.501, 0.128)
approxPost(-0.9266915, 20, 0.501, 0.128)
```

Q5
Bayes Factor: 27.3076
Post Prob: 0.9647
Mu: 0.5325 Sigma: 0.1401
```{r}
tval2(q5$PropCorOo3, q5$PropCorOc3)
marPred(2.882307, 25, 0.501, 0.192)/nullPred(2.882307, 25)

postProb(27.3076)

seeInference(2.882307, 25, 0.501, 0.192)
approxPost(2.882307, 25, 0.501, 0.192)
```

Q6
Bayes Factor:4663.3505
Post Prob: 0.9998
Mu: 0.8128 Sigma: 0.15615616
```{r}
tval2(q6$PropCorOo3, q6$PropCorOc3)
marPred(5.3797, 25, 0.637, 0.204)/nullPred(5.3797, 25)

postProb(4663.3505)

seeInference(5.3797, 25, 0.637, 0.204)
approxPost(5.3797, 25, 0.637, 0.204)
```

Q7
Bayes Factor: 164.425
Post Prob: 0.9940
Mu: 0.4845  Sigma: 0.1041
```{r}
tval2(q7$PropCorOo3, q7$PropCorOc3)
marPred(4.190279, 25, 0.396, 0.116)/nullPred(4.190279, 25)

postProb(164.425)

seeInference(4.190279, 25, 0.396, 0.116)
approxPost(4.190279, 25, 0.396, 0.116)
```

Q8
Bayes Factor: 0.7779
Post Prob: 0.4375
Mu: 0.1882 Sigma: 0.1001001
```{r}
tval2(q8$PropCorOo3, q8$PropCorOc3)
marPred(0.414595, 25, 0.228, 0.116)/nullPred(0.414595, 25)

postProb(0.7779)

seeInference(0.414595, 25, 0.228, 0.116)
approxPost(0.414595, 25, 0.228, 0.116)
```

